An Empirical Analysis of the Role of Energy in Economic Growth

### 2014-12-25 Articles/Article_EnerEcon_2014/EconGrowth3.Rnw
Edits to Energy Econ paper.

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ data).

Coefficients for each resampled CES model for all countries and all nests
are shown in 
Figures~\ref{fig:CES_lambda_gamma_graph}--\ref{fig:CES_sigma_delta_graph}. 
The white crosshair on each graph shows the location 
of the fit to historical data. 
In Figures~\ref{fig:CES_sigma1_delta1_graph} and~\ref{fig:CES_sigma_delta_graph},
the bottom of the vertical axis represents $\sigma = 0$.
The top of the vertical axis represents $\sigma = \infty$.

@@

\twocolumn{}
<<LINEXq_Resample_Graph, fig.pos="H", fig.align='center', fig.width=2, fig.height=7, fig.cap="LINEX model (with energy) resampling results.">>=
linexCoeffsData <- filter(coeffs_df, model=="linex", energy=="iQp")
standardScatterPlot(data=linexCoeffsData, aes(a_0, c_t), facet_formula=country~energy, alpha=0.05) +
    labs(x=expression(a[0]), y=expression(c[t])) +
     scale_x_continuous(limits=c(-1, 2), breaks=c(-1, 0, 1.0, 2.0)) +
     scale_y_continuous(limits=c(0, 10), breaks=c(0, 5, 10))
@

@@

@@

@@

We further caution against using $R^2$ only as the criteria for choosing a model, 
because, in our experience, 
both Cobb-Douglas and CES models yield high $R^2$ values for all countries. 
(See Figures~\ref{fig:CD_Spaghetti_Graph} and~\ref{fig:CES_Spaghetti_Graph}.) 
LinEx models fare poorly in some cases and might be excluded on that basis. 
(See SA and TZ in Figure~\ref{fig:LINEX_Spaghetti_Graph}.) 
It is more important to know the uncertainty of estimated parameters than to assess $R^2$. 
With rigorous uncertainty testing as we have done here, 
the quality of parameter estimates is exposed: 
at a glance, the tightness of resampled results indicates 
whether or not confidence in the model and data is warranted. 
We conclude that it is insufficient to argue 
in favor of a particular production function 
on the grounds of its ability to fit historical data.

@@

>>>>

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ time series).

Coefficients for each resampled CES model for all countries and all nests
are shown in 
Figures~\ref{fig:CES_lambda_gamma_graph}--\ref{fig:CES_sigma_delta_graph}. 
The white crosshair on each graph shows the location 
of the fit to historical data. 
In Figures~\ref{fig:CES_sigma1_delta1_graph} and~\ref{fig:CES_sigma_delta_graph},
the bottom of the vertical axis represents $\sigma = 0$ (perfect complements).
The top of the vertical axis represents $\sigma = \infty$ (perfect substitutes).

@@

\twocolumn{}
<<LINEXq_Resample_Graph, fig.pos="H", fig.align='center', fig.width=2, fig.height=7, fig.cap="LINEX model (with energy) resampling results.">>=
linexCoeffsData <- filter(coeffs_df, model=="linex", energy=="iQp")
standardScatterPlot(data=linexCoeffsData, aes(a_0, c_t), facet_formula=country~energy, alpha=0.05) +
    labs(x=expression(a[0]), y=expression(c[t])) +
     scale_x_continuous(limits=c(0, 2), breaks=c(0, 1.0, 2.0)) +
     scale_y_continuous(limits=c(0, 10), breaks=c(0, 5, 10))
@

@@

**** Need to think this presentation through, 
want to see how balances with the rest of the paper; 
do we have the space seems to be the limiting factor? --MdW ****

**** I'm fine if we want to change to another form of these tables.
But, I found the presentation below to be a good way to think about the results.
My idea was to say ``hang on! We have several ways to get at some of these elasticities.
Let's compare all the different ways to look at the elasticity between any two factors of production.
Do they agree?  
Are they in conflict?''  
Another point to make is that these comparisons need to be done honestly. 
We should not claim anything when the data are uncertain.

Regarding Martin's question about length, 
I've seen other papers in \emph{Energy Economics} that are 51 printed journal pages. (Saunders, 2008)
I expect we are fine on length. --MKH ****

@@

%++++++++++++++++++++++++++++++
\subsection{Which production function is best for energy augmentation?} 
\label{sec:which_best}
%++++++++++++++++++++++++++++++

**** Shall we have a section here discussing which function is ``best''? --MKH ****
What are the criteria?

Allows assymetric uncertainty estimation via resampling: all work fine.

Best match to historical data: CES.

No ambiguity from model structure: CD and LinEx OK here. 
CES has the undesireable characteristic that form (nesting) affects outcome.

Easy to estimate: CD and LinEx can be fitted with linear function. 
CES requires complex and time-consuming optimzation techniques.

Small uncertainties in fitted parameters: CD and LinEx better than CES.

Our assessment: None of the production functions is perfect. 
Probably CD is best compromise among ease of estimating, good (but not best) fit to historical data, and
small uncertainties in fitted parameters.

Caution: none of the production functions analyzed herein allows for complex energy-economy phenomena
like rebound effect, backfire, and super-conservation.

@@

**** Not sure whether contextualization will work. 
Could be read that macroeconomic growth modeling in such cases is not really useful at all? 
--MdW ****

We further caution against using $R^2$ only as the criteria for choosing a model, 
because, in our experience, 
both Cobb-Douglas and CES models yield high $R^2$ values for all countries. 
(See Figures~\ref{fig:CD_GDP_Spaghetti_Graph} and~\ref{fig:CES_Spaghetti_Graph}.) 
LinEx models fare poorly in some cases and might be excluded on that basis. 
(See SA and TZ in Figure~\ref{fig:LINEX_Spaghetti_Graph}.) 
It is more important to know the uncertainty of estimated parameters than to assess $R^2$. 
With rigorous uncertainty testing as we have done here, 
the quality of parameter estimates is exposed: 
at a glance, the tightness of resampled results indicates 
whether or not confidence in the model and data is warranted. 
We conclude that it is insufficient to argue 
in favor of a particular production function 
on the grounds of its ability to fit historical data.

@@

**** Other notes from our word document ****

Lastly, what do aggregate models say?

What direction can take from this work?

What is the status of ME growth modeling after this?
